How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-28 tse-paper/content.tex
Fix some blibs

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
The \emph{PROSE library}~\cite{prose2019webpage} is the basis of
our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to synthesize, PBE returns the
error message ``no program found.''

\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program.
% I hope this is fine, you can also cut this sentence if it still
% bothers you :)
If the program finds no match because the analyzed build logs
does not contain the structure defined by the training examples,
PBE returns an error message.

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique that
retrieve those lines from a build log which are most
similar to examples given by the user.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the search query.
The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the outputs of the training examples determines how many of
the most similar lines are returned as the output of the retrieval
run.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
We associate each example with keywords which appear in the targeted
chunk or close to it.

KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average of lines in the output of the training examples.

@@

To fully comprehend how difficult a task build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines randomly from the build log (RLR).
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples, thus arguably
giving it a slight advantage.

@@

To illustrate the chunk retrieval techniques further, we want to give
an example of real-world log LogChunks.
\Cref{lst:chunk-example} shows two excerpts
from Android build logs.
Marked in blue are the chunks
which contain the information on why the
corresponding build failed.

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
(which has a
different structural representation) to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

CTS will rank the lines from an analyzed build log which are most
similar to the chunks in the training examples
from \Cref{lst:chunk-example}.
It will then return the two lines ranked highest, as the chunks in
the training examples have on average two lines.

Keywords that appear close to the chunk within
the build log configure the technique KWS\@.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If we train KWS with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

@@

To investigate when the chunk retrieval techniques are suited to
analyze build logs (\textbf{RQ2}), we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular, we are interested in how many examples a technique
should be trained with (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}), and how reliable the quality of the
produced outputs is (\textbf{RQ2.3}).
This leads us to the following research questions:

@@

	% actually that is now giving recall another (nicer) name
	% improved if we just talk about recall later?
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}\vspace{0.1cm}$
	Partially Successful retrieval = $\mathit{true}\ \mathit{iff}\
	0 < \mathit{Recall} < 1, \mathit{false\ otherwise}\vspace{0.1cm}$
	Unsuccessful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 0, \mathit{false\ otherwise}\vspace{0.1cm}$

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
\label{fig:results-CTS}
\end{figure*}

% You might even say Figure 12-15 show two
% plots in the short text under the section heading.
% It just makes the
% text flow so much better.

@@

The first and most important aspect that influences a choice for
a chunk retrieval technique are the structural categories.
Are the targeted log chunks always presented in
the same structural way within the build logs? Then the information
chunks in all training examples and the analyzed build log are in the
same structural category.
This is why the first question in \Cref{fig:crt-recommendation} asks
how many structural categories the targeted log chunks have.

>>>>

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide in an automatically generated program.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
The \emph{PROSE library}~\cite{prose2019webpage} is the basis of
our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable \emph{PROSE} to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to be synthesized, PBE returns the
error message ``no program found.''

\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program.
% I hope this is fine, you can also cut this sentence if it still
% bothers you :)
If the program finds no match because the analyzed build logs
do not contain the structure the training examples define,
PBE returns an error message.

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique that
retrieves those lines from a build log which are most
similar to examples given by the user.

@@

To improve the similarity calculation, we prune very often or
very rarely appearing words.
Finally, the
% TODO TF-IDF undefined
algorithm weighs the vectors using TF-IDF, a best practice for natural
language queries~\cite{lee1997document}.

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the search query.
The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the outputs of the training examples determines how many of
the most similar lines our approach returns as the output of the retrieval
run.

@@

\subsubsection{Configuration}
A set of keywords configures the chunk retrieval with KWS\@.
To better
compare KWS with PBE and CTS, we also configure it through examples.
% TODO: unclear
We associate each example with keywords which appear in the targeted
chunk or close to it.

% TODO: unclear. or do you mean: same number == all searched for ?
KWS then searches for those keywords which are tied to the greatest
number of examples in the training set.
If several keywords are associated with the same number of training
examples, and no other keywords are associated with more training
examples, KWS searches for all of these keywords.

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average of lines in the specification output of the training examples.

@@

To fully comprehend how difficult a task build log analysis 
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines at random from the build log (RLR).
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples, thus arguably
giving it a slight advantage over guessing without previous knowledge.

@@

% TODO: want to?
% real-world log LogChunks ?
To illustrate the chunk retrieval techniques further, we want to give
an example of real-world log LogChunks.
\Cref{lst:chunk-example} shows two excerpts
from Android build logs.
Marked in blue are the chunks
which contain the information on why the
corresponding build failed.

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3}
(which has a
different structural representation) to the training set,
PBE is not able to synthesize a program anymore and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

CTS will rank the lines from an analyzed build log which are most
similar to the chunks in the training examples
from \Cref{lst:chunk-example}.
It will then return the two lines ranked highest, as the chunks in
the training examples have on average 2.0 lines.

Keywords that appear close to the chunk within
the build log configure the technique KWS\@.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed.''
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If we train KWS with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

@@

To investigate when the chunk retrieval techniques are suited to
analyze build logs, we perform an empirical comparison
study on the \emph{LogChunks} data set~\cite{brandt2020logchunks}.
In particular, we are interested in how many examples a technique
should be trained with (\textbf{RQ2.1}),
how similar the structural representation of the targeted log chunks
has to be (\textbf{RQ2.2}), and how reliable the quality of the
produced outputs is (\textbf{RQ2.3}).
This leads us to the following research questions:

@@

	% actually that is now giving recall another (nicer) name
	% improved if we just talk about recall later?
	\item Successful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 1, \mathit{false\ otherwise}\vspace{0.1cm}
        \\$
	Partially Successful retrieval = $\mathit{true}\ \mathit{iff}\
	0 < \mathit{Recall} < 1, \mathit{false\ otherwise}\vspace{0.1cm}
        \\$
	Unsuccessful retrieval = $\mathit{true}\ \mathit{iff}\
	\mathit{Recall} = 0, \mathit{false\ otherwise}\vspace{0.1cm}$

@@

\end{subfigure}\hspace{\fill}
\begin{subfigure}[tb]{\columnwidth}
		\centering
				\includegraphics[width=\columnwidth,
				clip]{img/big-study/recall-precision-categorycount-CTS.pdf}
		\caption{structural category count
		in training and test set.}
		\label{fig:recall-precision-categorycount-CTS}
\end{subfigure}
\caption{Precision, recall, and F$_{1}$-score of chunk
retrieval with Common Text Similarity (CTS) compared by \ldots}
% TODO: where are the bean plots?
\label{fig:results-CTS}
\end{figure*}

% TODO 
% You might even say Figure 12-15 show two
% plots in the short text under the section heading.
% It just makes the
% text flow so much better.

@@

The first and most important aspect that influences a choice for
a chunk retrieval technique are the structural categories. In other words,
are the targeted log chunks always presented in
the same structural way within the build logs? Then the information
chunks in all training examples and the analyzed build log are in the
same structural category.
This is why the first question in \Cref{fig:crt-recommendation} asks
how many structural categories the targeted log chunks have.
